A Simple Model of Energy Transition Under a Carbon Budget

### 2015-04-26 DraftApril2715.tex
Prelim Results

%
% File: thesisexample.tex   Version 1.8   May 6, 1999

%  1. To get appendices, you don't do anything different from a normal
%     report document.  That means, put the command \appendix before
%     you begin your first appendix, then do each appendix with a
%     \chapter command.  Note that if you have only one appendix, it is
%     customary to leave it unnumbered.  Do this with \chapter*.
%
%  2.  If you use \chapter*, which produces unnumbered chapters, you 
%       have to add that chapter to the table of contents by hand, e.g.
%
%         \chapter*{Appendix}
%         \addcontentsline{toc}{chapter}{Appendix}
%

%  4.  Problems with math formulas in chapter headings:
% 
%         a.  Any lowercase letters in the formula are converted to
%         uppercase, e.g. f(x) becomes F(X).   If you really need
%         lowercase math letters in your chapter titles, use the
%         option plainchapterheads (and, if you want, type your
%         chapter titles in ALL CAPS so that the appearance doesn't
%         change).  Note there is no problem 
%         for section or subsection headings in either case.  (Options
%         such as plainchapterheads are given as part of the 
%         \documentclass command, see below under ``Document Options'').
%
%         b.  Some perfectly reasonable math commands when used in
%         \chapter give the error
%          ``LATEX ERROR: \command  ALLOWED ONLY IN MATH MODE.''
%         The solution to this is to do
%        
%              \newcommand{\mymath}{problem math goes here}
%       
%         and then
%        
%              \chapter{All about \protect\mymath}
%
%         also, the option plainchapterheads will fix this too.
%
%  5. If your References section doesn't show up in the table of
%     contents, you need to add the line \addcontentsline...
%     as done at end of this file.   Make sure that you include the 
%     page break command as done there or else you may end up
%     with the wrong page number in your table of contents.

%\documentclass{easychithesis}

% Document Options: 
%
% Note if you want to save paper when printing drafts, replace the
% above line by
% 
\documentclass[singlespace]{easychithesis}
% 
%  Similarly, if you need to use the plainchapterheads option, you do
%   
%  \documentclass[plainchapterheads]{easychithesis}
%
% You can give more than one option, if you desire.
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\newcommand{\Lagr}{\mathcal{L}}

@@

% Create the official title
\title{A Simple Model of Energy Transition Under a Carbon Budget} 
\author{Adriana Ciccone}
\date{\today}
\department{Environmental Science \& Policy}
\division{The Harris School of Public Policy Studies} 
\degree{Master of Science} 
\maketitle

@@

%\topmatter{Acknowledgements}
%I'd like to thank Jason Bermack, who unknowingly played
%\emph{Big Generator} for me in his car, when we were both in high 
%school, thus ensuring a long and fruitful relationship with Yes music 
%throughout my college days.

%I'd also like to thank Dave Moulton, who likes Yes too, and 
%accompanied me to my first ever rock concert - Yes in the Oakland 
%Coliseum.  They played ``Awaken'' for 20 minutes.  Wow.

@@

%
% List of figures
%

@@

%
% Body Chapters
%
\chapter{Introduction}
\section{Relevance}

@@

Power generation is accomplished by dollars of capital ($k^h(t), k^l(t)$) that can produce certain amounts of electricity in a given year.Capital is comprised of an initial investment $H_0$ or $L_0$ that depreciates exponentially, but can be supplemented with positive investments $H^+(t)$ or $L^+(t)$ that also depreciate exponentially, or retired early with $H^-(t)$ or $L^-(t)$.

Power is generated by multiplying each type of capital in a year by the overnight capital cost scaled by a capacity factor ($F^h(t), F^l(t)$. High-emitting (dirty) capital has a lower capital cost than low-emitting (clean) capital, but also has higher emissions ($m^h(t) > m^l(t) \forall t$).

@@

Note that the relationship between capital (the stocks) and investments (the flows) can be represented by the following equations:

\begin{equation}
k^h(t) = H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx
\end{equation}

\begin{equation}
k^l(t) = L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx
\end{equation}

@@

\begin{equation}
G = \bar{F}^h\left (H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \right) + \bar{F}^l(1-e^{-t/\tau}) \left ( L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx\right )
\end{equation}

@@

@@

@@

\begin{equation}
\frac{d\Lagr}{d\mu} = \bar{E} - mF^h \left [(1-e^{-N/n})(1-n)H_0 n + n^2(N + n(1-e^{-N/n})) + n\int_0^N H^+(t) dt - \int_0^N\int_0^t H^-(x)dx dt   \right] \geq 0
\end{equation}

@@

@@

>>>>

%  1. To get appendices, you don't do anything different from a normal
%     report document.  That means, put the command \appendix before
%     you begin your first appendix, then do each appendix with a
%     \chapter command.  Note that if you have only one appendix, it is
%     customary to leave it unnumbered.  Do this with \chapter*.

%  5. If your References section doesn't show up in the table of
%     contents, you need to add the line \addcontentsline...
%     as done at end of this file.   Make sure that you include the 
%     page break command as done there or else you may end up
%     with the wrong page number in your table of contents.

\documentclass[singlespace]{easychithesis}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\newcommand{\Lagr}{\mathcal{L}}
\usepackage{tabulary}

@@

\title{A Simple Model of Energy Transition Under a Carbon Budget} 
\author{Adriana Ciccone}
\date{\today}
\department{Environmental Science \& Policy}
\division{The Harris School of Public Policy Studies} 
\degree{Master of Science} 
\maketitle

@@

%\topmatter{Acknowledgements}

@@

@@

\chapter{Introduction}
\section{Relevance}

@@

Power generation is accomplished by dollars of capital ($k^h(t), k^l(t)$) that can produce certain amounts of electricity in a given year.Capital is comprised of an initial investment $H_0$ or $L_0$ that depreciates exponentially, but can be supplemented with positive investments $H^+(t)$ or $L^+(t)$ that also depreciate exponentially, or retired early with $H^-(t)$ or $L^-(t)$.  \\

Power is generated by multiplying each type of capital in a year by the overnight capital cost scaled by a capacity factor ($F^h(t), F^l(t)$. High-emitting (dirty) capital has a lower capital cost than low-emitting (clean) capital, but also has higher emissions ($m^h(t) > m^l(t) \forall t$). \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Math}

\paragraph{} The region or country of interest produces electricity by two methods: clean (low emitting, represented by $k^l(t)$) and dirty (high emitting, represented by $k^h(t)$). Clean capital produces electricity with low carbon dioxide emissions, but, as a newer technology, it is more expensive per unit of power output than dirty. Since it is a new technology, the capital cost is expected to decrease over time as the production process matures and technological innovation produces lower-cost units. Dirty capital is a mature technology that is cheap to build but emits a higher level of carbon dioxide per unit of electricity produced than clean. Since it is a mature technology, there is little reason to believe that production costs will decline substantially in the future.

\paragraph{} Each type of capital depreciates exponentially over time. The decision variables for the optimization are additional investments in each type of capital that also depreciate at the same rate, represented by four functions ($H^+(t), H^-(t), L^+(t), \& L^-(t)$). These decision variables define the optimal path of investment in dirty (high emitting) and clean (low emitting) capital. All four functions are non-negative over their domain. The functions $H^+(t)$ and $L^+(t)$ represent additional dollars of investment to build new capital of each type. $H^-(t)$ and $L^-(t)$ remove capital from the economy via early retirement. There is no scrap cost to early retirement - the economy simply loses the ability to produce electricity from that particular facility. As I show in the next section, this has consequences for the ability of the economy to meet it's total electricity demand. The retired plant cannot be re-started. Any need for additional capital in time t must be met by a positive investment in either $H^+(t)$ or $L^+(t)$.

\begin{equation}
k^h(t) = H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \geq 0 \qquad \forall t
\end{equation}

\begin{equation}
k^l(t) = L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx \geq 0 \qquad \forall t
\end{equation}

These equations express total capital in each time period as a function of some exponential depreciation and all prior investments. At each time t, the initial capital endowment, $H_0 = H^+(0)$ and $L_0 = L^+(0)$ depreciate exponentially and any further positive investments depreciate based on their age (x-t). The integral in the above expression shows how each investment $H^+(t)$ or $L^+(t)$ is weighted by an exponential term of the difference in that investment's age and the current time period. Therefore we see that early investments have a more negative exponential term and count less towards existing capital than more recent investments. While we have no limits on the amount of capital in each time period other than the non-negativity constraint on the decision variables, a physical limitation remains that we cannot have negative amounts of capital. We therefore restrict both $k^h(t)$ and $k^l(t)$ to be non-negative in all time periods.

\paragraph{} The objective is to minimize total additional positive investment in both high and low carbon intensity capital for the lowest cost transition to clean energy. We minimize over positive investment only to avoid negative investment - early retirement - benefiting the objective. Instead, early retirement forces additional investment later in the simulation due to an energy generation constraint. That is, if we retire a certain amount of capital early and have a shortfall in generation capacity, we must make a positive investment at some later point. This extra generation would be counted as additional positive investment in the below objective function, penalizing early retirement. The sum is discounted by an interest rate so that we minimize over the net present value of investments.

\begin{equation}
C(H^+(t), L^+(t)) = \int_0^N (H^+(t) + L^+(t))e^{-rt}) dt \quad if H(t), L(t) \geq 0
\end{equation}

%% No operating cost, so when we early retire, we leave the pile of coal that we had purchased upfront - biases the result up/down? Clean doesn't have this problem %%

\paragraph{} The economy has some total energy demand that can follow any exogenous path over time. This represents total annual electricity demand, in billions of kilowatt-hours, and therefore daily and seasonal fluctuations in power demand are ignored. Capital must be used to produce electricity to meet this demand, and we must convert from dollars of capital to kilowatt-hours of electricity to take in to account the relative cost difference of producing one unit of clean energy versus one unit of dirty energy.

\begin{equation}
G(t) = F^h(t) k^h(t) + F^l(t) k^l(t)
\end{equation}

In this equation, $F^h (t)$ and $F^l (t)$ are the overnight capital cost per billion kilowatt-hours per year. That is, they represent the number of kilowatt-hours per year that each dollar of capital can produce. In early years when the clean technology is still relatively new and expensive, a dollar of capital may not produce much energy per year, but a dollar of clean capital invested later in the model may produce significantly more energy. It is important to be clear however, on what this cost is. It is not an annual maintenance cost or any measure of financial obligation for additional production. It is simply the the ability of that unit (dollar) of capital to produce a certain amount of electricity.

%% We don't track vintages over time, so these costs are just an assumed exogenous trajectory. Which bias? Also we take care of this in the vintaged version%%

\paragraph{} An emissions cap is negotiated by the region's government. Over the course of a fixed number of years, the region must not emit more than a negotiated percent of business as usual emissions. In practice, the government could negotiate to agree to a limit on total number of tons of $CO_2$ emitted instead of negotiating a percentage of business as usual. However, it is convenient to represent a bulk emissions cap as a percent reduction for both modeling and to understand how our optimal cost will change as a function of total allowed emissions.

\begin{equation}
E_{max} = \alpha \int_0^N m^h(t) F^h(t) H^+(0) + m^l(t) F^l(t) L^+(0) dt \qquad \alpha \in [0,1]
\end{equation}

To arrive at an emissions cap, we transform dollars of capital to kilowatt-hours per year as above, then multiply by the carbon intensity of energy, $m^h(t)$ and $m^l(t)$. This value is a measurement of the pounds of $CO_2$ emitted from power generation per kilowatt-hour. We can then write actual emissions over the length of simulation as the following.

\begin{equation}
E(H, L) = \int_0^N  m^h(t) F^h(t) H_0 + m^l(t) F^l(t) L_0 dt
\end{equation}

Total emissions over the treaty's length must not exceed the cap represented by a fraction of business as usual, leading to the final constraint:

\begin{equation}
E(H,L) \leq E_{max}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solving the Model Analytically}

In order to attempt an analytical solution, I make several key assumptions about the functional forms of key parameters. While the equations for capital stocks remain the same, I assume a flat, fixed energy demand, $G$, through the length of the simulation. The energy per unit of capital for high-emitting technologies is similarly fixed at $F^h$ for the entire period, while it decays exponentially for low-emitting technology with a half life of $\tau$. Assuming these functional forms is arguably a good approximation of existing beliefs about the future costs of low-emitting technologies with learning by doing or economies of scale; that is, a rapid improvement in the output per dollar of capital that slowly levels off as t grows large. I further assume that high-emitting carbon intensity is fixed at $m^h$ and that there are no emissions associated with clean capital. With these simplification we now write the following optimization problem. \\

@@

\begin{multline}
G = \bar{F}^h\left (H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \right) \\ + \bar{F}^l(1-e^{-t/\tau}) \left ( L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx\right )
\end{multline}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Note that a solution will go here if I can get one. For now, the first order conditions are as follows:

@@

\begin{multline}
\frac{d\Lagr}{d\mu} = \bar{E} - mF^h \left [(1-e^{-N/n})(1-n)H_0 n + n^2(N + n(1-e^{-N/n})) \right.\\ + \left. n\int_0^N H^+(t) dt - \int_0^N\int_0^t H^-(x)dx dt   \right] \geq 0
\end{multline}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Model Parameters}
\subsection{Functional Form Choices}

\subsection{Parameter Values}

\begin{tabulary}{1.0\textwidth}{LLLL}
    Variable & Description & Value & Source\\
    \hline
alpha &  Regional Land Area  &  Not Specified  &  N/A \\
area0 &  Initial Regional Land Area  &  Not Specified  &  N/A \\
landloss &  Regional Land Loss due to Sea Level Rise  &  Not Specified  &  N/A \\
	\hline
\end{tabulary}

\subsection{Implementation}
link to source code

@@

\includegraphics[width=0.9\textwidth]{Fl_Simulation_Range.png}

\includegraphics[width=0.9\textwidth]{cap_and_invest_results_A_1.png}
\includegraphics[width=0.9\textwidth]{cap_and_invest_results_A_25.png}

\includegraphics[width=0.9\textwidth]{cap_and_invest_results_A_50.png}

\includegraphics[width=0.9\textwidth]{cap_and_invest_results_A_75.png}
\includegraphics[width=0.9\textwidth]{cap_and_invest_results_A_100.png}

\includegraphics[width=0.9\textwidth]{cap_and_invest_results_Fl_1.png}
\includegraphics[width=0.9\textwidth]{cap_and_invest_results_Fl_25.png}
\includegraphics[width=0.9\textwidth]{cap_and_invest_results_Fl_50.png}
\includegraphics[width=0.9\textwidth]{cap_and_invest_results_Fl_75.png}
\includegraphics[width=0.9\textwidth]{cap_and_invest_results_Fl_100.png}

\includegraphics[width=0.9\textwidth]{cost_as_func_of_Fl_emissions_05.png}

\includegraphics[width=0.9\textwidth]{cost_as_func_of_Alpha.png}
